AbxD01 Analysis

### 2015-02-23 abxD01_analysis.Rmd
Clean comments and set evals to TRUE for rendered file

This figure shows the correlation analysis of bacterial species present on Day 0 with *C. difficile* levels on Day 1.

@@

```{r, eval=FALSE}

@@

results = NULL
results <- matrix(c(otu, cor.spear, pval.spear, cor.ken, pval.ken), ncol=5)
colnames(results) <- c('otu','corSpear','pvalSpear', "corKen", "pvalKen")
results <- results[order(results[,3]),]  #order by pvalue column=3
write.table(results[1:dim(results)[1],], file="~/Documents/Github/abxD01/Figure 2/abxD01.final.an.unique_list.0.03.subsample.filter16mintotal.shared.correl.topdose2.results.txt", sep="\t", row.names=FALSE)

@@

This figure shows correlation values for OTUs with next day *C. difficile* CFU levels calculated across the original set of antibiotic experiments.

@@

The code for this is the script "stacked.barcharts.r" which contains the stackedbarcharts() function in the paper's home Github directory. Opening this file, the parameters I changed in the top were:
```{r, eval=FALSE}
setwd("/Users/alyxschubert/Documents/Github/abxD01/")

abx<-c("5 mg/ml", "0.5 mg/ml", "0.1 mg/ml")
stackedbarcharts(file = "./Figure 3/abxD01.final.tx.2.subsample.allceftitr.forlogscale.csv", fileIDS = "./Figure 3/allceftitr_tx2_barchart_ids.csv", graphLabels = abx, sortbyphyl = TRUE, graphbyphyl = TRUE, excludeGroups = "1untrCef")

abx<-c("5 mg/ml", "0.5 mg/ml", "0.1 mg/ml")
stackedbarcharts(file = "./Figure 3/abxD01.final.tx.2.subsample.allstreptitr.forlogscale.csv", fileIDS = "./Figure 3/allstreptitr_tx2_barchart_ids.csv", graphLabels = abx, sortbyphyl = TRUE, graphbyphyl = TRUE, excludeGroups = "1untrStrep")

abx<-c("5 mg/ml", "0.5 mg/ml", "0.1 mg/ml")
stackedbarcharts(file = "./Figure 3/abxD01.final.tx.2.subsample.allvanctitr.forlogscale.csv", fileIDS = "./Figure 3/allvanctitr_tx2_barchart_ids.csv", graphLabels = abx, sortbyphyl = TRUE, graphbyphyl = TRUE, excludeGroups = "1untrVanc")

@@

The file/fileIDS should be matched by the titration.
Input files on github:
1. abxD01.final.tx.2.subsample.allceftitr.forlogscale.csv
2. abxD01.final.tx.2.subsample.allstreptitr.forlogscale.csv
3. abxD01.final.tx.2.subsample.allvanctitr.forlogscale.csv
4. allceftitr_tx2_barchart_ids.csv
5. allstreptitr_tx2_barchart_ids.csv
6. allvanctitr_tx2_barchart_ids.csv

For this graph I decided to take out the untreated communities to highlight the differences among titration groups. So in r I changed the file a bit:

```{r, eval=FALSE}
file<-file[file$expgroup!="1untrStrep",]
#alternatively for the other titrated antibiotics it's "1untrCef", or "1untrVanc"
```

@@

For comparison of the CFU's among titration groups: 
```{r, eval=TRUE}
file<-read.csv("~/Documents/Github/abxD01/Figure 3/abxD01.final.tx.2.subsample.allstreptitr.forlogscale.csv", header=T)
file<-file[file$expgroup!="1untrStrep",]
avgCD<-tapply(file$nextDayCFU, file$expgroup, mean)
strep_avgCD<-format(avgCD, scientific=TRUE, digits=2)
file$nextDayCFU <- log10(file$nextDayCFU+1)
#kruskal.test(file$nextDayCFU ~ file$expgroup)
# p < 0.0001
#strep_CFU_wilcox <- pairwise.wilcox.test(file$nextDayCFU, file$expgroup, p.adj="BH")
# all significantly different
strep.aov<-aov(file$nextDayCFU ~ file$expgroup)
anova(strep.aov)
#TukeyHSD(strep.aov)
strep_ttest <- pairwise.t.test(file$nextDayCFU, file$expgroup, p.adjust="BH")
print(strep_ttest)
# all significantly different

file<-read.csv("~/Documents/Github/abxD01/Figure 3/abxD01.final.tx.2.subsample.allceftitr.forlogscale.csv", header=T)
file<-file[file$expgroup!="1untrCef",]
avgCD<-tapply(file$nextDayCFU, file$expgroup, mean)
cef_avgCD<-format(avgCD, scientific=TRUE, digits=2)
file$nextDayCFU <- log10(file$nextDayCFU+1)
#kruskal.test(file$nextDayCFU ~ file$expgroup)
# p = 0.003
#cef_CFU_wilcox <- pairwise.wilcox.test(file$nextDayCFU, file$expgroup, p.adj="BH")
# mid and low dose not different, but high different than mid and low
cef.aov<-aov(file$nextDayCFU ~ file$expgroup)
anova(cef.aov)
#TukeyHSD(cef.aov)
cef_ttest <- pairwise.t.test(file$nextDayCFU, file$expgroup, p.adjust="BH")
print(cef_ttest)
# all significantly different

file<-read.csv("~/Documents/Github/abxD01/Figure 3/abxD01.final.tx.2.subsample.allvanctitr.forlogscale.csv", header=T)
file<-file[file$expgroup!="1untrVanc",]
avgCD<-tapply(file$nextDayCFU, file$expgroup, mean)
vanc_avgCD<-format(avgCD, scientific=TRUE, digits=2)
file$nextDayCFU <- log10(file$nextDayCFU+1)
#kruskal.test(file$nextDayCFU ~ file$expgroup)
# p = 0.0016
#vanc_CFU_wilcox <- pairwise.wilcox.test(file$nextDayCFU, file$expgroup, p.adj="BH")
# mid and low dose not different, but high different than mid and low
vanc.aov<-aov(file$nextDayCFU ~ file$expgroup)
anova(vanc.aov)
#TukeyHSD(vanc.aov)
vanc_ttest <- pairwise.t.test(file$nextDayCFU, file$expgroup, p.adjust="BH")
print(vanc_ttest)
# mid and low dose not different, but high different than mid and low

For comparisons between family-level phylotypes:
Performed inside abxD01.barcharts.xOTU.test.r
```{r, eval=FALSE}
file<-read.csv("~/Documents/Github/abxD01/Figure 3/abxD01.final.tx.2.subsample.allstreptitr.forlogscale.csv", header=T)
file<-file[file$expgroup!="1untrStrep",]
kruskal.test(file$Otu001 ~ file$expgroup)
kruskal.test(file$Otu003 ~ file$expgroup)
kruskal.test(file$Otu002 ~ file$expgroup)
kruskal.test(file$Otu005 ~ file$expgroup)
kruskal.test(file$Otu004 ~ file$expgroup)
kruskal.test(file$Otu014 ~ file$expgroup)
kruskal.test(file$Otu006 ~ file$expgroup)
kruskal.test(file$Otu007 ~ file$expgroup)
kruskal.test(file$Otu009 ~ file$expgroup)
kruskal.test(file$Otu015 ~ file$expgroup)
kruskal.test(file$Otu017 ~ file$expgroup)
kruskal.test(file$Other ~ file$expgroup)

test <- pairwise.wilcox.test(file$Otu001, file$expgroup, p.adj="BH")
comparisons <- length(test$p.value[]) - sum(is.na(test$p.value))

pairwise.wilcox.test(file$Otu003, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu005, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu002, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu004, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu014, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu006, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu007, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu009, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu015, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu017, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Other, file$expgroup, p.adj="BH")

file<-read.csv("~/Documents/Github/abxD01/Figure 3/abxD01.final.tx.2.subsample.allvanctitr.forlogscale.csv", header=T)
file<-file[file$expgroup!="1untrVanc",]
kruskal.test(file$Otu001 ~ file$expgroup)
kruskal.test(file$Otu002 ~ file$expgroup)
kruskal.test(file$Otu003 ~ file$expgroup)
kruskal.test(file$Otu005 ~ file$expgroup)
kruskal.test(file$Otu004 ~ file$expgroup)
kruskal.test(file$Otu014 ~ file$expgroup)
kruskal.test(file$Otu006 ~ file$expgroup)
kruskal.test(file$Otu007 ~ file$expgroup)
kruskal.test(file$Otu008 ~ file$expgroup)
kruskal.test(file$Otu019 ~ file$expgroup)
kruskal.test(file$Otu018 ~ file$expgroup)
kruskal.test(file$Otu017 ~ file$expgroup)
kruskal.test(file$Other ~ file$expgroup)

pairwise.wilcox.test(file$Otu001, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu003, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu005, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu002, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu004, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu014, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu006, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu007, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu008, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu019, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu018, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu017, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Other, file$expgroup, p.adj="BH")

file<-read.csv("~/Documents/Github/abxD01/Figure 3/abxD01.final.tx.2.subsample.allceftitr.forlogscale.csv", header=T)
file<-file[file$expgroup!="1untrCef",]
kruskal.test(file$Otu001 ~ file$expgroup)
kruskal.test(file$Otu002 ~ file$expgroup)
kruskal.test(file$Otu003 ~ file$expgroup)
kruskal.test(file$Otu005 ~ file$expgroup)
kruskal.test(file$Otu004 ~ file$expgroup)
kruskal.test(file$Otu014 ~ file$expgroup)
kruskal.test(file$Otu006 ~ file$expgroup)
kruskal.test(file$Otu007 ~ file$expgroup)
kruskal.test(file$Otu008 ~ file$expgroup)
kruskal.test(file$Otu009 ~ file$expgroup)
kruskal.test(file$Otu019 ~ file$expgroup)
kruskal.test(file$Otu018 ~ file$expgroup)
kruskal.test(file$Otu016 ~ file$expgroup)
kruskal.test(file$Otu024 ~ file$expgroup)
kruskal.test(file$Otu011 ~ file$expgroup)
kruskal.test(file$Other ~ file$expgroup)

pairwise.wilcox.test(file$Otu001, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu003, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu005, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu002, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu004, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu014, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu006, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu007, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu008, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu009, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu011, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu019, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu018, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu016, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Otu024, file$expgroup, p.adj="BH")
pairwise.wilcox.test(file$Other, file$expgroup, p.adj="BH")

@@

#what are the important variables (via permutation) #type 1 is mean decrease in accuracy, type 2 is mean decrease in node impurity
par(mfrow=c(1, 1)) 
varImpPlot(model_tx2, type=1)
imp<-importance(model_tx2)
most.imp<-rownames(imp[order(imp[, "%IncMSE"], decreasing=T)[1:17],])
```

>>>>

This figure shows the correlation analysis of bacterial species present on Day 0 with *C. difficile* levels on Day 1 using the original set of antibiotic groups.

@@

```{r, eval=TRUE}

@@

results = NULL
results <- matrix(c(otu, cor.spear, pval.spear, cor.ken, pval.ken), ncol=5)
colnames(results) <- c('otu','corSpear','pvalSpear', "corKen", "pvalKen")
results <- results[order(results[,3]),]  #order by pvalue column=3
#write.table(results[1:dim(results)[1],], file="~/Documents/Github/abxD01/Figure 2/abxD01.final.an.unique_list.0.03.subsample.filter16mintotal.shared.correl.topdose2.results.txt", sep="\t", row.names=FALSE)
print(results[1:dim(results)[1],])

@@

@@

The file/fileIDS should be matched by the titration.
Input files on github:
1. abxD01.final.tx.2.subsample.alltitrations.forlogscale.csv
2. alltitrations_tx2_barchart_ids.csv

Uses functions in R scripts folder: stacked.barcharts.r, findStatLetters.r, barplotBeside.r

```{r, eval=TRUE}

stackedbarcharts(file = "~/Documents/Github/abxD01/Figure 3/abxD01.final.tx2.subsample.alltitrations.forlogscale.csv", fileIDS = "~/Documents/Github/abxD01/Figure 3/alltitrations_tx2_barchart_ids.csv", divide = TRUE)

@@

@@

For comparison of the CFU's among titration groups: 
```{r, eval=TRUE}
file<-read.csv("~/Documents/Github/abxD01/Figure 3/abxD01.final.tx2.subsample.alltitrations.forlogscale.csv", header=T)
file<-file[!grepl("untr", file$expgroup),]
file$nextDayCFU <- log10(file$nextDayCFU+1)

cef <- file[file$abx=="Cefoperazone", 3:4]
strep <- file[file$abx=="Streptomycin", 3:4]
vanc <- file[file$abx=="Vancomycin", 3:4]

diffs <- function(data){
  data.aov<-aov(data$nextDayCFU ~ data$expgroup)
  anova(data.aov)
  pairwise.t.test(data$nextDayCFU, data$expgroup, p.adjust="BH")
}

diffs(cef)
diffs(strep)
diffs(vanc)

@@

#what are the important variables (via permutation) #type 1 is mean decrease in accuracy, type 2 is mean decrease in node impurity
par(mfrow=c(1, 1)) 
varImpPlot(model_tx2, type=1)
#imp<-importance(model_tx2)
#most.imp<-rownames(imp[order(imp[, "%IncMSE"], decreasing=T)[1:17],])
```
